---
title: "Statistical assignment 1"
author: "Jessica Ledger 660013603"
date: "29/1/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Open data (10 points)

In this assignment you will work with the individual level data from wave 8 of the Understanding Society survey. First, you need to open the data set. Please complete the code below.

```{r}
library(tidyverse)
library(dplyr)

Data <- read_tsv("/Users/jessicaledger/Desktop/WORK/THIRD YEAR/Data Analysis III/data2020/data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab")
```

Now you have got your data frame stored as Data. 

## Select variables (10 points)

The data for Wave 8 of the Understanding Society were collected in 2016-18. Among other things, people were asked the following question: "Should the United Kingdom remain a member of the European Union or leave the European Union?" In this assignment, we will explore how answers to this question depend on sex and age.

First, you need to select the variables for the analysis. You want to keep the following variables: cross-wave individual identifier (*pidp*), support for the UK remaining or leaving the EU (*h_eumem*), sex (*h_sex_dv*), age (*h_age_dv*), and sample origin (*h_memorig*).

Complete the code below to select those variables from the data frame and save the result.

```{r}
Data <- Data %>%
        select(pidp, h_eumem, h_sex_dv, h_age_dv, h_memorig)
```

## Filter observations (10 points)

To make nationally representative estimates from the Understanding Society data we would need to use weight coefficients. There are many different types of weight coefficients that can be used depending on the question and the level of analysis (see the User Guide, pp. 65-71). We will not do this in this assignment. However, what we want to do is to keep data from the original Understanding Society sample only (ukhls gb 2009-10), dropping  data for Northern Ireland, the BHPS cohort members and ethnic minority boost samples. This will make data closer to be representative for Great Britain. You need to choose the observations where *h_memorig* has the value of 1.

```{r}
Data <- Data %>%
        filter(h_memorig == 1)
```

## Recode data (20 points)

Let us tabulate the variables for EU support, sex, and age.

```{r}
table(Data$h_eumem)
table(Data$h_sex_dv)
table(Data$h_age_dv)

```

You will see that all these variables are numeric. You can learn what the numeric codes mean by checking the codebook here: https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/datafile/indresp/wave/8

We want to do the following:

1) Recode the variable for EU support as binary (1 for Remain, 0 for Leave), coding all types of missing values (including refusals and "don't know") as NA.

2) Recode sex into a character vector with the values "male" or "female".

3) Recode age into a variable with the following categories: 16 to 25, 26 to 40, 41 to 55, 56 to 70, over 70.

In all cases, we want to create new variables.

```{r}

Data <- Data %>%
        mutate(EU = ifelse(h_eumem == "1", "1",
                           ifelse(h_eumem == "2", "0", NA_real_))
      
        ) %>%
        mutate(sex = recode(h_sex_dv,
                            "1" = "male",
                            "2" = "female",
                            .default = NA_character_)) %>%
        mutate(agegr = case_when(
                between(h_age_dv, 16, 25) ~ "16 to 25",
                between(h_age_dv, 26, 40) ~ "26 to 40",
                between(h_age_dv, 41, 55) ~ "41 to 55",
                between(h_age_dv, 56, 70) ~ "56 to 70",
                h_age_dv >= 70 ~ "over 70"
        )
        )

class(Data$EU)
class(Data$sex)
class(Data$agegr)


```

## Summarise data (20 points)

Let us **dplyr** to calculate how many people in the sample supported Remain and Leave, both as absolute numbers and percentages.

```{r}


Data %>%
        count(EU) %>%
  mutate(perc = n / sum(n) * 100)

        
```

Write a couple of sentences with the interpretation of this result. How this compares with the result of the 2016 referendum? Why?

This is an interesting result as there is not an overall majority for either Remain or EU (neither Remain or Leave recieved 50% of the vote). The result of the sample shows that 48% voted to Remain in the EU and only 41% voted to Leave the EU. This is very different in comparison to what happened overall in the 2016 referendum as in 2016, 46.6% voted in favour of Remain and 53.4% voted in favour of Leave. This given sample would indicate that more people were in favour to Remain in the EU, yet this is not what happened in 2016. The difference in result may be as a result of a smaller sample group (approx. 11% of sample have been removed as NA) or as this was a questionnaire taken before the referndum vote took place and does not determine what people actually voted, therefore people could have changed their votes closer to when they cast their vote.

## Summarise data by sex and age (30 points)

Now let us look at the support for Leave and Remain by sex and age. Use your newly created variables.


```{r}
Data %>%
  group_by(sex) %>%
  count(EU) %>%
  mutate(percentage = n / sum(n) * 100)

Data %>%
  group_by(agegr) %>%
  count(EU) %>%
  mutate(percentage = n/sum(n) * 100)

Data %>%
  group_by(sex, agegr) %>%
  count(EU) %>%
  mutate(percentage = n/sum(n) * 100)
        
```


Write a couple of sentences interpreting your results.

Regarding the support for Leave and Remain by sex, a higher proportion of women voted to remain (51%) in the EU rather than leave (40%). Similarly, a higher proportion of men voted remain (45%) rather than leave (43%). It is clear that there is a closer gap between the vote for men (2%) and there is a wider gap with women (11%.)

Regarding the support for Leave and Remain by age, it can be interpreted that a larger proportion of young people voted to remain in the EU (61% 16 to 25, and 56% 26 to 40) and the middle ages were closer in their divide between leaving and remaining in the EU. Furthermore, older people, most noteably over 70 were much more likely to be pro Leave rather than remaining in the EU (50% voted Leave in over 70 category)


